### 2017-11-03 submission/manuscript.Rmd
update results

@@

**Background.**

**Results.**

@@

***Genera Previously Associated with Carcinoma Increases Relative Risk More than Alpha Diversity:***
Both fecal and tissue samples had a significantly increased RR for carcinoma but not for adenoma [Figure 3] which was greater than either evenness or Shannon diversity [Figure 2 & 3]. The relative risk did not increase when considering the total abundance or increasing number of carcinoma associated genera [Figure 3]. The RR effect size was greater for stool (RR range = `r paste(round(min(filter(select_RR_stool, !(measure %in% c("all_four", "total_four")))[, "rr"]), digits = 2))` - `r paste(round(max(filter(select_RR_stool, !(measure %in% c("all_four", "total_four")))[, "rr"]), digits = 2))`) then for tissue (RR range = `r paste(round(min(filter(select_RR_tissue, !(measure %in% c("all_four", "total_four")))[, "rr"]), digits = 2))` - `r paste(round(max(filter(select_RR_tissue, !(measure %in% c("all_four", "total_four")))[, "rr"]), digits = 2))`). This decrease may be explained by the fact that tissue samples include matched samples.

>>>>

adn_select_RR_stool <- as.data.frame(read_csv("../data/process/tables/adn_select_genus_RR_stool_composite.csv"))
adn_select_RR_tissue <- as.data.frame(read_csv("../data/process/tables/adn_select_genus_RR_tissue_composite.csv"))

@@

**Background.** An increasing body of literature suggests that there a role for the microbiota in colorectal cancer. Important players within this axis have ranged from individual microbes to the whole community. A recent meta-analysis investigated this but only focused on potential biomarkers. This study expands on this previous research and tests the hypothesis that detectable changes in the bacterial community are important both to increasing relative risk and model accuracy for adenoma and carcinoma. To test this hypothesis we examined both fecal and tissue samples from 2109 individuals across 14 different studies.

**Results.** There was a significant decrease from control to adenoma to carcinoma for both Shannon diversity and evenness for fecal samples (P-value `r ifelse(min(filter(stool_lme, !(alpha_metric %in% "sobs"))[, "BH"]) < 0.05, invisible("<"), invisible("check tables"))` 0.05) after correcting for study and variable region. Lower Shannon diversity and evennes in fecal samples resulted in a significant increase in relative risk for carinoma (P-value `r ifelse(min(filter(alpha_RR_crc_stool, !(measure %in% "sobs"))[, "pvalue"]) < 0.05, invisible("<"), invisible("check tables"))` 0.05) but not adenoma (P-value `r ifelse(min(filter(alpha_RR_adn_stool, !(measure %in% "sobs"))[, "pvalue"]) > 0.05, invisible(">"), invisible("check tables"))` 0.05). Previously associated colorectal cancer genera (*Fusobacterium*, *Parvimonas*, *Peptostreptococcus*, or *Porphyromonas*) followed a similar pattern with a significantly increased relative risk by their presence for carcinoma (P-value `r ifelse(min(filter(select_RR_stool, !(measure %in% c("all_four", "total_four")))[, "pvalue"]) < 0.05, invisible("<"), invisible("check tables"))` 0.05) but not adenoma (P-value `r ifelse(min(filter(adn_select_RR_stool, !(measure %in% c("all_four", "total_four", "Porphyromonas")))[, "pvalue"]) > 0.05, invisible(">"), invisible("check tables"))` 0.05) with the exception of *Porphyromonas* (P-value `r ifelse(min(filter(adn_select_RR_stool, measure == "Porphyromonas")[, "pvalue"]) < 0.05, invisible("<"), invisible("check tables"))` 0.05).

@@

***Genera Previously Associated with Carcinoma Increases Relative Risk More than Alpha Diversity:***
Both fecal and tissue samples had a significantly increased RR for carcinoma but not for adenoma [Figure 3]. The relative risk for feces was greater than either evenness or Shannon diversity [Figure 2 & 3]. The relative risk did not increase when considering the total abundance or increasing number of carcinoma associated genera [Figure 3]. The RR effect size was greater for stool (RR range = `r paste(round(min(filter(select_RR_stool, !(measure %in% c("all_four", "total_four")))[, "rr"]), digits = 2))` - `r paste(round(max(filter(select_RR_stool, !(measure %in% c("all_four", "total_four")))[, "rr"]), digits = 2))`) then for tissue (RR range = `r paste(round(min(filter(select_RR_tissue, !(measure %in% c("all_four", "total_four")))[, "rr"]), digits = 2))` - `r paste(round(max(filter(select_RR_tissue, !(measure %in% c("all_four", "total_four")))[, "rr"]), digits = 2))`). This decrease may be explained by the fact that tissue samples include matched samples.

Only two measures for adenoma were significant when investigating these CRC associated genera. The first was *Porphyromonas* for feces (RR range: `r round(filter(adn_select_RR_stool, measure == "Porphyromonas")[, "ci_lb"], digits = 2)` - `r round(filter(adn_select_RR_stool, measure == "Porphyromonas")[, "ci_ub"], digits = 2)`, P-value = `r format(filter(adn_select_RR_stool, measure == "Porphyromonas")[, "pvalue"], digits = 2, nsmall = 2)`) and the second was the total relative abundance of all four CRC associated genera on tissue (RR range: `r round(filter(adn_select_RR_tissue, measure == "total_four")[, "ci_lb"], digits = 2)` - `r round(filter(adn_select_RR_tissue, measure == "total_four")[, "ci_ub"], digits = 2)`, P-value = `r format(filter(adn_select_RR_tissue, measure == "total_four")[, "pvalue"], digits = 3, nsmall = 3)`).

### 2017-11-03 submission/manuscript.tex
update results

\textbf{Background.}

\textbf{Results.}

@@

\textbf{\emph{Genera Previously Associated with Carcinoma Increases
Relative Risk More than Alpha Diversity:}} Both fecal and tissue samples
had a significantly increased RR for carcinoma but not for adenoma
{[}Figure 3{]} which was greater than either evenness or Shannon
diversity {[}Figure 2 \& 3{]}. The relative risk did not increase when
considering the total abundance or increasing number of carcinoma
associated genera {[}Figure 3{]}. The RR effect size was greater for
stool (RR range = 1.78 - 2.64) then for tissue (RR range = 1.33 - 1.53).
This decrease may be explained by the fact that tissue samples include
matched samples.

>>>>

\textbf{Background.} An increasing body of literature suggests that
there a role for the microbiota in colorectal cancer. Important players
within this axis have ranged from individual microbes to the whole
community. A recent meta-analysis investigated this but only focused on
potential biomarkers. This study expands on this previous research and
tests the hypothesis that detectable changes in the bacterial community
are important both to increasing relative risk and model accuracy for
adenoma and carcinoma. To test this hypothesis we examined both fecal
and tissue samples from 2109 individuals across 14 different studies.

\textbf{Results.} There was a significant decrease from control to
adenoma to carcinoma for both Shannon diversity and evenness for fecal
samples (P-value \textless{} 0.05) after correcting for study and
variable region. Lower Shannon diversity and evennes in fecal samples
resulted in a significant increase in relative risk for carinoma
(P-value \textless{} 0.05) but not adenoma (P-value \textgreater{}
0.05). Previously associated colorectal cancer genera
(\emph{Fusobacterium}, \emph{Parvimonas}, \emph{Peptostreptococcus}, or
\emph{Porphyromonas}) followed a similar pattern with a significantly
increased relative risk by their presence for carcinoma (P-value
\textless{} 0.05) but not adenoma (P-value \textgreater{} 0.05) with the
exception of \emph{Porphyromonas} (P-value \textless{} 0.05).

@@

\textbf{\emph{Genera Previously Associated with Carcinoma Increases
Relative Risk More than Alpha Diversity:}} Both fecal and tissue samples
had a significantly increased RR for carcinoma but not for adenoma
{[}Figure 3{]}. The relative risk for feces was greater than either
evenness or Shannon diversity {[}Figure 2 \& 3{]}. The relative risk did
not increase when considering the total abundance or increasing number
of carcinoma associated genera {[}Figure 3{]}. The RR effect size was
greater for stool (RR range = 1.78 - 2.64) then for tissue (RR range =
1.33 - 1.53). This decrease may be explained by the fact that tissue
samples include matched samples.

Only two measures for adenoma were significant when investigating these
CRC associated genera. The first was \emph{Porphyromonas} for feces (RR
range: 1.06 - 2.29, P-value = 0.023) and the second was the total
relative abundance of all four CRC associated genera on tissue (RR
range: 1.21 - 3.89, P-value = 0.00933)
